Topological Equivariant Artist Model

### 2022-03-29 slides/main.tex
flipped arrows

\begin{document}

\title{Topological Equivariant Artist Model}

@@

\section{Introduction}
\begin{frame}<presentation:1|handout:0>{What are we doing?}
    \begin{itemize}
        \item develop a model for describing data to graphic transformations
        \item specify a visualization library architecture based on this model
        \item implement functional(ish) components based on this model using ideas from functional programming
    \end{itemize}
\end{frame}

@@

\begin{frame}<presentation:1|handout:0>{Case Study: Matplotlib}
    \begin{figure}
       \includegraphics[width=\linewidth]{figures/flow/artists.png}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Visualizations Preserve Structure}
    \begin{description}
        \item [continuity] topological properties \cite{wilkinsonGrammarGraphics2005}, i.e. how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
        \item [equivariance] data and visual encodings are matched such that transformations have an equivalent effect on data and graphical representations, e.g. rotating a matrix and image, shifting  points on a line and a line graphic
      \end{description}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Continuity}
    \begin{figure}
        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Expressed in container} 
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Why is container type not enough?}
    \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
\end{frame}

\begin{frame}<presentation:1-|handout:0>{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
        \pause
        \item[Expressiveness] which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
    \end{description}
\end{frame}

%flip columns
\begin{frame}<presentation:1-|handout:0>{Frameworks for Expressing Visual Equivariance}
    \begin{description}
        \item[linguistically] visualization has syntax, semantics, grammar expresses how to design structure preserving visualizations
        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}

        \item[algebraically] transformations on data and graphics are equivalent and symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
        \begin{columns}
            \column{.5\textwidth}
            \begin{description}
                \item[D] raw data 
                \item[R] representation (data container) 
                \item[V] visualization
            \end{description}
            \column{.5\textwidth}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
                \end{tikzcd}
                \end{equation*}
        \end{columns} 
    \item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
    \end{description}
\end{frame}

\begin{frame}<presentation:1-|handout:1>
    \frametitle{Domain specific libraries know their structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\        
        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}<presentation:1-|handout:1>
    \frametitle{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.3\textheight]{../paper/figures/k_different_types.png}
    \end{figure}

@@

\section{Algebraic Topology & Category Theory}
\begin{frame}<presentation:1|handout:1>{Mathematical Data Abstraction}
    \begin{description}
        \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
        \item[Category Theory Language] express constraints in specifications \cite{wielsManagementEvolvingSpecifications1998}   
        \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
    \end{description}
\end{frame}

@@

               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
               \onslide*<7|handout:4>{\arrow[u, "\gsectionc \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}"', bend right, pos=.5]}

@@

    \pause 
    \only<5|handout:1>{

@@

                \item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}

@@

\begin{frame}<presentation:1-|handout:1>{Category $\textcolor{source}{\mathcal{C}}$}
\begin{columns}
    \column{0.39\textwidth}
    \begin{equation*}
    \begin{tikzcd}[ampersand replacement=\&]
        \onslide*<1-|handout:1>{\textcolor{source}{C_1}} 
        \onslide*<3-|handout:1>{\arrow[r, "f", color=source]}
        \onslide*<4-|handout:1>{\arrow[rd, "g\circ f"', color=source]}
        \onslide*<2-|handout:1>{\arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source]} \& 
        \onslide*<1-|handout:1>{\textcolor{source}{C_2}}
        \onslide*<3-|handout:1>{\arrow[d, "g", color=source]} 
        \onslide*<2-|handout:1>{\arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]} \\\& 
        \onslide*<1-|handout:1>{\textcolor{source}{C_3}}
        \onslide*<2-|handout:1>{\arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]}              
    \end{tikzcd}
    \end{equation*}
    \column{0.59\textwidth}
    \begin{alertblock}<5-|handout:1>{associativity} 
        if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
    \end{alertblock}
    \begin{alertblock}<6-|handout:1>{identity} 
        for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
    \end{alertblock}
    \end{columns}
\end{frame}

\begin{frame}<presentation:1-|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
    \begin{columns}
        \column{0.49\textwidth}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \textcolor{fade}{C_1} \arrow[r, "f", color=fade] \arrow[rd, "g\circ f"', color=fade] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=fade] \& \textcolor{fade}{C_2} \arrow[d, "g", color=fade] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=fade] \\
        \& \textcolor{fade}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=fade]              
        \end{tikzcd}
        \end{equation*}
        \column{0.49\textwidth}
        \onslide*<presentation:2|handout:1>{
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&]
            \textcolor{source}{C_1} \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[l, "f"',color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]\\
            \& \textcolor{source}{C_3} \arrow[u, "g"', color=source] \arrow[lu, "f \circ g", color=source] \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]
            \end{tikzcd}
        \end{equation*}
        }
        \end{columns}
\end{frame}

@@

\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&]
            \dtotalc_a 
            \arrow[dd, "\dfunc c_a"', color=action] \& \& 
            \dtotalc_a \times \dtotalc_b 
            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
            \arrow[ll, "\pi_a"', color=total] 
            \arrow[rr, "\pi_b", color=total] \& \& 
            \dtotalc_b 
            \arrow[dd, "\dfuncc_b", color=action] \\
            \& \& \& \& \\
            \dtotalc_a  \& \& 
            \dtotalc_a \times \dtotalc_b 
            \arrow[rr, "\pi_b", color=total] 
            \arrow[ll, "\pi_a"', color=total] \& \& 
            \dtotalc_b
        \end{tikzcd}
    \end{equation*}    
\end{frame}

@@

@@

\begin{frame}<presentation:1|handout:1>
    \frametitle{Transformations in Data $\vchannelc: \dsectionc \mapsto \vsectionc$}
    %rewrite in gamma notation
    \begin{equation*}
    \begin{tikzcd}[ampersand replacement = \&]
        \dtotalc 
        \arrow[dd, "\vchannelc_{\opensetc}"', color=artist] 
        \arrow[rr, "\dfuncpullc"] 
        \& \& 
        \dfuncpullc\dtotalc 
        \arrow[rr, "\dfunctc_{\dtotalc}"] 
        \&  \& 
        \dtotalc^{\prime} 
        \arrow[dd, "{\vchannelc^{\prime}}_{\opensetc^{\prime}}", color=artist] \\
        \& 
        \opensetc 
        \arrow[lu, dotted, maps to] 
        \arrow[ld, dotted, maps to] \& 
        \opensetc^{\prime} 
        \arrow[l, "\dfunchc"] 
        \arrow[u, dotted, maps to] 
        \arrow[d, dotted, maps to] 
        \arrow[rrd, dotted, maps to] 
        \arrow[rru, dotted, maps to] 
        \& \& \\
        \vtotalc 
        \arrow[rr, "\dfuncpullc"'] 
        \& \& 
        \dfuncpullc\vtotalc 
        \arrow[rr, "\dfunctc_{\vtotalc}"'] 
        \& \& 
        \vtotalc^{\prime}
    \end{tikzcd}
    \end{equation*}
    Given $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$ and  $\vsectionc(\dbasepointc) = \vsectionc(\dfunchc(\dbasepointc^{\prime}))$
    \begin{description}
        \item[equivariance] $\dfunctc_{\vtotalc}(\vchannelc_{\opensetc}(\dsectionc))
        = \vchannelc_{\opensetc}(\dfunctc_{\dtotalc}(\dsectionc))$, $\vchannelc_{\opensetc} = {\vchannelc^{\prime}}_{\opensetc^{\prime}}$
        \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
    \end{description}
\end{frame}

\begin{frame}{Data to Visual Transformation $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
    %remove maps to 
$\pi(\dtotalc) = \pi(\vchannelc(\dtotalc))$ and $\vchannelc$ is composable s.t 
    \begin{columns}
        \column{.5\textwidth}
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiberc_{\dbasepointc} 
            \arrow[rr, "\vchannelc", maps to] 
            \arrow[dd, "\dfunctc_{\dtotalc}"'] 
            \& \& 
            \vfiberc_{\dbasepointc} 
            \arrow[dd, "\dfunctc_{\vtotalc}"] \\
            {} 
            \arrow[rr, dashed, maps to]
            \& \& {} \\
            {\dfiberc^{\prime}}_{\dbasepointc} 
            \arrow[rr, "\vchannelc^{\prime}"', maps to]
            \& \& 
            {\vfiberc^{\prime}}_{\dbasepointc}
            \end{tikzcd}
        \end{equation*}
        \column{.5\textwidth} 
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
            \dfiberc_{\dbasepointc} 
            \arrow[rr, "\vchannelc_{\dtotalc} ", maps to, color=artist] 
            \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, maps to, color=artist] 
            \&  \& 
            \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} \arrow[dd, "\vchannelc_{\vtotalc}", maps to, color=artist] \\
            \& \& \\
            \&  \& {\vfiberc_{\dbasepointc}}^{\vtotalc}
            \end{tikzcd}
        \end{equation*}
    \end{columns}

@@

\begin{frame}<presentation:1|handout:1>
    \frametitle{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiberc_a\times \dfiberc_b 
            \arrow[rr, "\vchannelc", color=artist, maps to] 
            \arrow[rd, "p_{\dfiberc}", color=fiber] \& \& \vfiberc_a\times \vfiberc_b 
            \arrow[rd, "p_{\vfiberc}", color=fiber] \& \\    \& 
            \dfiberc_{a} 
            \arrow[rr, "\vchannelc_{a}", dashed, maps to, color=artist] \& \& \vfiberc_a \\
            \dfiberc_a\times \dfiberc_c 
            \arrow[rr, "\vchannelc^{\prime}", color=artist, maps to] 
            \arrow[ru, "p_{\dfiberc^{\prime}}"', color=fiber] \& \& 
            {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} 
            \arrow[ru, "p_{\vfiberc^{\prime}}"', color=fiber] \&           
            \end{tikzcd}
    \end{equation}
    \begin{alertblock}{Consistent Transformations \cite{hullmanKeeping2018}}
        if $p_{\dfiberc} = p_{\dfiberc^{\prime}}$ then $p_{\vfiberc}(\vchannelc(\dsectionc)) = p_{\vfiberc^{\prime}}(\vchannelc^{\prime}(\dsectionc^{\prime}))$ s.t. there exists a transformation $\vchannelc_a: \dfiberc_{a}\rightarrow \vfiberc_{a}$
    \end{alertblock}
\end{frame}

>>>>

\begin{document}
\begin{frame}<presentation:1-|handout:0>{About}
    Slides: https://bit.ly/team329
    \begin{block}{Me (twitter/github: story645)}
        \begin{itemize}
            \item nth year grad student (on my 3rd EO)
            \item former adjunct at CCNY, former Digital Fellow
            \item Matplotlib Community Manager \& Core Developer
        \end{itemize}
    \end{block}

    \begin{block}<presentation:2|handout:0>{Project}
        \begin{itemize}
            \item Funded by Chan Zuckerberg Initiative EOSS 1 \& 3
            \item paper rejected by vizweek last spring
            \item work has since gone all in on category theory 
        \end{itemize}
    \end{block}
\end{frame}

@@

\section{Introduction}
\begin{frame}<presentation:0|handout:0>{What are we doing?}
    \begin{itemize}
        \item develop a model for describing data to graphic transformations
        \item specify a visualization library architecture based on this model
        \item implement functional(ish) components based on this model using ideas from functional programming
    \end{itemize}
\end{frame}

@@

\begin{frame}<presentation:1|handout:0>{Matplotlib}
    \begin{figure}
       \includegraphics[width=\linewidth]{figures/flow/artists.png}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1-|handout:0>
    \frametitle<presentation:1-2|handout:0>{Everything is an Artist}
    \frametitle<presentation:3|handout:0>{Transformations Change Artists into Different Coordinates}
    \begin{overprint}
        \onslide<1|handout:0>\includegraphics[scale=.25]{figures/math/sphx_glr_anatomy_001_2_0x.png}
        \onslide<2|handout:0>\includegraphics[width=\textwidth]{figures/math/artist_tree.png}
        \onslide<3|handout:0>\includegraphics[width=\textwidth]{figures/math/transforms.png}
    \end{overprint}
\end{frame}

\begin{frame}[fragile]{A Simple Artist}
\begin{minted}{python}
    class SomeArtist(Artist):
        'An example Artist that implements the draw method'

        def draw(self, renderer):
            # create some objects and use renderer to draw self here
            renderer.draw_path(graphics_context, path, transform)
\end{minted}
\end{frame}

\begin{frame}<presentation:1-|handout:0>{Goals}
    \begin{figure}
        \includegraphics[width=\linewidth]{figures/flow/artists.png}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1|handout:0>{How do we express structure?} 
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Continuity}
    \begin{figure}
        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
    \end{figure}
    \begin{block}{Topological Properties \cite{wilkinsonGrammarGraphics2005}}
        how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
    \end{block}
\end{frame}

\begin{frame}<presentation:1|handout:0>{Visual Algorithms \& Continuity}
    \begin{figure}
        \includegraphics[width=1\linewidth]{figures/math/plottypes.png}
    \end{figure}
\end{frame}

\begin{frame}<presentation:1-|handout:0>{Equivariance}
    \begin{description}
        \item[What] Retinal Variables \& Marks: visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Why] Graphical Integrity: graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Why] Naturalness: easier to understand when properties match\cite{norman_things_smart}
        \pause
        \item[How] Expressiveness: which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
    \end{description}
\end{frame}

%flip columns
\begin{frame}<presentation:0-|handout:0>{Theoretical Frameworks}
    \begin{description}
        \item[linguistically] visualization has syntax, semantics, grammar expresses how to design structure preserving visualizations
        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}

        \item[algebraically] transformations on data and graphics are equivalent and symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
        \begin{columns}
            \column{.5\textwidth}
            \begin{description}
                \item[D] raw data 
                \item[R] representation (data container) 
                \item[V] visualization
            \end{description}
            \column{.5\textwidth}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
                \end{tikzcd}
            \end{equation*}
        \end{columns} 
    %\item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
    \end{description}
\end{frame}

\begin{frame}<presentation:1-|handout:1>
    \frametitle{Domain Specific Library: library assumes structure \cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\        
        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}<presentation:1-|handout:1>
    \frametitle{Building Block Library\cite{wongsuphasawatNavigatingWideWorld2021}: visual algorithms assume structure \cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.3\textheight]{../paper/figures/k_different_types.png}
    \end{figure}

@@

\section{Algebraic Topology & Category Theory}
\begin{frame}<presentation:1|handout:1>{Design Composable Structure Preserving API}
    \begin{description}
        \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
        \item[Category Theory Language] express constraints in specifications \cite{wielsManagementEvolvingSpecifications1998}   
        \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
    \end{description}
\end{frame}

@@

               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc"', bend right, pos=.5]}
               \onslide*<7|handout:4>{\arrow[u, "\gsectionc"', bend right, pos=.5]}

@@

    \only<4-5|handout:1>{

@@

                \only<5:handout:1>{\item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}}
    \only<4:handout:1>{

@@

\begin{frame}<presentation:1-|handout:1->
    \frametitle<presentation:1-6|handout:1>{Category \setc}
    \frametitle<presentation:7|handout:2>{Category \setc^{op}}
\begin{columns}
    \column{0.39\textwidth}
    \begin{equation*}
    \begin{tikzcd}[ampersand replacement=\&]
        \onslide*<1-|handout:1>{\setc_1} 
        \onslide*<3-6|handout:1>{\arrow[r, "r_{1,2}", color=set]}
        \onslide*<4-6|handout:1>{\arrow[rd, "r_{1,2}\circ r_{2,3}"', color=set]}
        \onslide*<2-6|handout:1>{\arrow["id_{\setc_1}"', loop, distance=2em, in=215, out=145, color=set]} \& 
        \onslide*<1-|handout:1>{\setc_2}
        \onslide*<7|handout:2>{\arrow[l, color=set]}
        \onslide*<3-6|handout:1>{\arrow[d, "r_{2,3}", color=set]} 
        \onslide*<2-6|handout:1>{\arrow["id_{\setc_2}"', loop, distance=2em, in=125, out=55, color=set]} \\\& 
        \onslide*<1-|handout:1>{\setc_3}
        \onslide*<7|handout:2>{\arrow[u, color=set]}
        \onslide*<7|handout:2>{\arrow[lu, color=set]}
        \onslide*<2-6|handout:1>{\arrow["id_{\setc_3}"', loop, distance=2em, in=305, out=235, color=set]}              
    \end{tikzcd}
    \end{equation*}
    \column{0.59\textwidth}
    \begin{alertblock}<5-6|handout:1>{associativity} 
        if $r_{1,2}: \setc_1 \rightarrow \setc_2$, $r_{2,3}: \setc_2 \rightarrow \setc_3$ and $r_{3,4}: \setc_3 \rightarrow \setc_4$ then $r_{3,4}\circ (r_{2,3} \circ r_{1,2}) = (r_{3,4} \circ r_{2,3}) \circ r_{1,2}$
    \end{alertblock}
    \begin{alertblock}<6|handout:1>{identity} 
        for every $r_{1,2}: \setc_1 \rightarrow \setc_2$ there exists identity morphisms $r_{1,2} \circ id_{\setc_1} = r_{1,2} = id_{\setc_2} \circ r_{1,2}$
    \end{alertblock}
    \end{columns}
\end{frame}

\begin{frame}<presentation:0|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
    \begin{columns}
        \column{0.49\textwidth}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \textcolor{fade}{C_1} \arrow[r, "f", color=fade] \arrow[rd, "g\circ f"', color=fade] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=fade] \& \textcolor{fade}{C_2} \arrow[d, "g", color=fade] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=fade] \\
        \& \textcolor{fade}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=fade]              
        \end{tikzcd}
        \end{equation*}
        \column{0.49\textwidth}
        \onslide*<presentation:2|handout:1>{
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&]
            \textcolor{source}{C_1} \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[l, "f"',color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]\\
            \& \textcolor{source}{C_3} \arrow[u, "g"', color=source] \arrow[lu, "f \circ g", color=source] \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]
            \end{tikzcd}
        \end{equation*}
        }
        \end{columns}
\end{frame}

@@

@@

\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&]
            \dtotalc_a 
            \arrow[dd, "\dfunc a"', color=action] \& \& 
            \dtotalc_a \times \dtotalc_b 
            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
            \arrow[ll, "\pi_a"', color=total] 
            \arrow[rr, "\pi_b", color=total] \& \& 
            \dtotalc_b 
            \arrow[dd, "\dfuncc_b", color=action] \\
            \& \& \& \& \\
            \dtotalc_a  \& \& 
            \dtotalc_a \times \dtotalc_b 
            \arrow[rr, "\pi_b", color=total] 
            \arrow[ll, "\pi_a"', color=total] \& \& 
            \dtotalc_b
        \end{tikzcd}
    \end{equation*}
    if there exists functions $\dfuncc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfuncc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t. $\pi_a \circ \dfuncc_a = \dfuncc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfuncc_b = \dfuncc_{a,b} \circ \pi_b$ then $\dfuncc_{a,b} = (\dfuncc_a, \dfuncc_b)$

@@

\begin{frame}<presentation:1|handout:1>
    \frametitle{Transformations in Data $\vchannelc: \dsectionc \mapsto \vsectionc$}
    %rewrite in gamma notation
    \begin{figure}
        \includegraphics[.35\textwidth]{figures/math/nu_table.png}
    \end{figure}
    Given $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$ and  $\vsectionc(\dbasepointc) = \vsectionc(\dfunchc(\dbasepointc^{\prime}))$
    \begin{description}
        \item[equivariance] $\dfunctc_{\vtotalc}(\vchannelc_{\opensetc}(\dsectionc))
        = \vchannelc_{\opensetc}(\dfunctc_{\dtotalc}(\dsectionc))$, $\vchannelc_{\opensetc} = {\vchannelc^{\prime}}_{\opensetc^{\prime}}$
        \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
    \end{description}
\end{frame}

\begin{frame}{Data to Visual Transformation $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
    %remove maps to 
$\pi(\dtotalc) = \pi(\vchannelc(\dtotalc))$ and $\vchannelc$ is composable s.t 
    \begin{columns}
        \column{.15\textwidth}
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiberc_{\dbasepointc} 
            \arrow[rr, "\vchannelc"] 
            \arrow[dd, "\dfunctc_{\dtotalc}"'] 
            \& \& 
            \vfiberc_{\dbasepointc} 
            \arrow[dd, "\dfunctc_{\vtotalc}"] \\
            {} 
            \arrow[rr, dashed]
            \& \& {} \\
            {\dfiberc^{\prime}}_{\dbasepointc} 
            \arrow[rr, "\vchannelc^{\prime}"']
            \& \& 
            {\vfiberc^{\prime}}_{\dbasepointc}
            \end{tikzcd}
        \end{equation*}
        \column{.5\textwidth} 
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
            \dfiberc_{\dbasepointc} 
            \arrow[rr, "\vchannelc_{\dtotalc} ", color=artist] 
            \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, color=artist] 
            \&  \& 
            \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} 
            \arrow[dd, "\vchannelc_{\vtotalc}", color=artist] \\
            \& \& \\
            \&  \& {\vfiberc_{\dbasepointc}}^{\vtotalc}
            \end{tikzcd}
        \end{equation*}
    \end{columns}

@@

\begin{frame}<presentation:1|handout:1>
    \frametitle{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiberc_a\times \dfiberc_b 
            \arrow[rr, "\vchannelc", color=artist] 
            \arrow[rd, "p_{\dfiberc}", color=fiber] \& \& \vfiberc_a\times \vfiberc_b 
            \arrow[rd, "p_{\vfiberc}", color=fiber] \& \\    \& 
            \dfiberc_{a} 
            \arrow[rr, "\vchannelc_{a}", dashed, color=artist] \& \& \vfiberc_a \\
            \dfiberc_a\times \dfiberc_c 
            \arrow[rr, "\vchannelc^{\prime}", color=artist] 
            \arrow[ru, "p_{\dfiberc^{\prime}}"', color=fiber] \& \& 
            {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} 
            \arrow[ru, "p_{\vfiberc^{\prime}}"', color=fiber] \&           
            \end{tikzcd}
    \end{equation}
    \begin{alertblock}{Consistent Transformations \cite{hullmanKeeping2018}}
        if $p_{\dfiberc} = p_{\dfiberc^{\prime}}$ then $p_{\vfiberc}(\vchannelc(\dsectionc)) = p_{\vfiberc^{\prime}}(\vchannelc^{\prime}(\dsectionc^{\prime}))$ s.t. there exists a transformation $\vchannelc_a: \dfiberc_{a}\rightarrow \vfiberc_{a}$
    \end{alertblock}
\end{frame}
